The Hydrogen Epoch of Reionization Array Dish: Characterization with Electromagnetic Simulations

### 2015-10-26 DishSimulations/paper/DishSimulationsPaper_v0.tex

\section{The Impact of Reflections on Delay-Transform Power Spectra}\label{sec:Formalism}
\section{Electromagnetic Simulations of the HERA dish element}\label{sec:Simulations}
\section{Comparing Simulations Results to Measurements}
\label{sec:Comparison}
\section{The Effect of the HERA dish Chromaticity on Foreground Leakage and Sensitivity}
\label{sec:Foregrounds}
\section{Conclusions}
\label{sec:Conclusion}
\bibliographystyle{apj}
\bibliography{biblio}
\end{document}

>>>>

\section{The Impact of Reflections on Delay-Transform Power Spectra}\label{sec:Formalism}
In this section, we show how reflections in the analogue signal path of an antenna lead to foreground contamination of the EoR window. HERA's primary antenna element consists of a sleeved dipole element suspended 5\,m above the focus of a 14\,m diameter parabolic dish (Fig.~\ref{fig:Dish}).

\begin{figure}
\includegraphics[width=.5\textwidth]{figures/DishSA.png}
\caption{The HERA primary antenna element-one of 19 undergoing currently taking the first observations in the Karoo in South Africa. The antenna consists of a sleeved dipole suspended within a 2\,m diameter skirt, five meters above the ground at the focal point of a 14\,m diameter dish.}\label{fig:Dish}
\end{figure}

 We show in Appendix~\ref{app:Reflections} that if an astronomical radio signal with time dependence at the location of the feed, $s({\bf \hat{k}},t)$, experience reflections within the dish such that the voltage recorded in the feed is 
\begin{equation}
v(t) = \sum_n r_n({\bf \hat{k}}) s({\bf \hat{k}},t - \tau_n)
\end{equation}
Than the resulting visibilities obtained by cross correlating antenna $i$ and antenna $j$ are given by 
\begin{equation}
V'_{ij}(f) = \sum_{m,n} \int d \Omega R_{mn}({\bf \hat{k}}) e^{2 \pi i \Delta \tau_{mn}f} e^{2 \pi i {\bf u}_{ij} \cdot {\bf \hat{k}}} I(f,{\bf \hat{s}}),
\end{equation}
where $R_{mn}({\bf \hat{s}})=r_m r_n^*$, $I(f,{\bf \hat{s}})$ is the intensity of the signal on the sky, and can be decomposed into foregrounds and signal
\begin{equation}
I(f,{\bf \hat{s}}) = I_{fg}(f,{\bf \hat{s}})+I_{21}(f,{\bf \hat{s}})
\end{equation}
While $I_{fg}$ is on the order of $10^6$ times larger than $I_{21}$, the continuum emission from foregrounds is expected to be spectrally smooth. Hence, a {\it delay transform} defined by
\begin{equation}
\widetilde{V}_{ij}(\tau) = \int d f e^{-2 \pi i \tau f} V_{ij}(f)
\end{equation}
should, in principle, separate the foregrounds from the signal with foregrounds occupying the wedge. However, the introduction of the chromatic $e^{2 \pi i \Delta \tau_{mn}}$ terms leads to foreground contamination being convolved out to higher delays.
\begin{equation}
\widetilde{V}_{ij}'(\tau) = \sum_{m,n} \widetilde{V}_{ij,FG}^{mn}(\tau - \Delta \tau_{mn}) + \widetilde{V}_{ij,21}^{mn}(\tau - \Delta \tau_{mn})
\end{equation}
where $V_{ij}^{mn}$ is a visibility resulting from the effective primary beam, $R_{mn}({\bf \hat{k}})$, at delay $\Delta \tau_{mn}$ induced by the reflections.
\begin{equation}
V_{ij}^{mn} = \int d \Omega R_{mn}({\bf \hat{s}}) I(f,{\bf \hat{s}})e^{2 \pi i {\bf u}_{ij} \cdot {\bf \hat{s}}}
\end{equation}
The delay transform power spectrum estimate, $\widehat{P}$, is proportional to $|\widetilde{V}_{ij}'|^2$, hence 
\begin{align}\label{eq:Reflections}
\widehat{P} & \propto |\widetilde{V}_{ij,FG}(\tau)|^2 + |\widetilde{V}_{ij,21}(\tau)|^2 + 2 \text{Re}(\widetilde{V}_{ij,21}(\tau)\widetilde{V}_{ij,FG}^*(\tau)) \nonumber \\
&+ \sum_{|\Delta \tau_{mn}| > 0} 2 \text{Re} \left[   \widetilde{V}^*_{ij} (\tau) \widetilde{V}^{mn}_{ij} (\tau - \Delta \tau_{mn})  \right] \nonumber \\
&+ \sum_{\substack{|\Delta \tau_{op}|>0\\|\Delta \tau_{mn}|>0}} \widetilde{V}^{mn}_{ij} (\tau - \Delta \tau_{mn}) \widetilde{V}^{op*}_{ij}(\tau - \Delta \tau_{op})
\end{align}
In the absence of reflections, only the top line of equation~\ref{eq:Reflections} would enter the delay power spectrum estimate. The following lines involve the translation of the signal (including foregrounds and signal) out to the delay of each reflection and weighted by the effective primary beam of the reflection. Since each translated visibility is intrinsically $\sim 10^6$ times brighter than the signal, any reflection with a delay large enough to place it within the EoR window, must be sufficiently attenuated by $R_{mn}({\bf \hat{k}})$. The precise degree of attenuation is determined in \citep{Thyagarajan:2015c} using simulations of foregrounds and noise and sets the spec for our antenna. In this paper, we study $R_{mn}({\bf \hat{k}})$ for the HERA antenna element to verify that it both meets the spec and agrees with direct measurements of the antenna element with refelctometry \citep{Patra:2015} and Orbcomm measurements \citep{Neben:2015b}.

\section{Electromagnetic Simulations of the HERA dish element}\label{sec:Simulations}
In Fig.~\ref{fig:SimulationSetup} we show the geometery of the electromagnetic simulation. {\bf Rich: fill in the details here}
\begin{figure}

\subsection{The Simulations}
\includegraphics[width=.5\textwidth]{figures/One_dish_Pfeed_render_pw_0deg.png}
\caption{A rendering of our time domain simulation at $t=0$, demonstrating the geometry and setup of our electromagnetic simulation. The plane wave is started just above the feed (red plane).}\label{fig:SimulationSetup}
\end{figure}

\begin{figure*}
\includegraphics[width=\textwidth]{figures/SimulationIllustration.png}
\caption{An illustration of our simulation products and their origin in the HERA antenna geometry. A plane wave is injected from above the feed (red line). The amplitude of the electric field of the plane wave at output of the feed along with the voltage at the feed terminal outputs is recorded (black line). The feed in our simulation is situated $5$\,m above the bottom of the dish, hence there is a $\approx 30$\,ns delay between when the plane wave passes the terminal for the first time (A) and when it is first absorbed in the dipole (B), leading to the voltage response. Of concern to 21\,cm experiments are the subsequent reflections between the feed and the dish (C) which can lead to large delay contamination of the EoR window.}
\label{fig:Simulation}
\end{figure*}
\subsection{Deconvolving the Response Function}
In our simulation, we obtain the voltage at the feed output as a function of time which we will call $v_{out}(t)$. It is related to the input lane wave through the equation 
\begin{equation}
v_{out}(t) = \sum_n r_n(\theta=0) v_{in}(t - \tau_n)
\end{equation}
which is essentially a convolution in time of $r_n(\theta=0)$ with the input plane wave 
\section{Comparing Simulations Results to Measurements}
\label{sec:Comparison}
\section{The Effect of the HERA dish Chromaticity on Foreground Leakage and Sensitivity}
\label{sec:Foregrounds}
\section{Conclusions}
\label{sec:Conclusion}
\bibliographystyle{apj}
\bibliography{biblio}

\appendix
\section{The Effect of Reflections and Cross-Talk on Visibilities}\label{app:Reflections}
In this section, we develope formalism to discuss the impact of reflections of electromagnetic waves between antennas and within the signal chain of single antennas on foreground leakage in 21\,cm experiments. We start with the time varying electric field from a single source with location ${\bf \widehat{k}}$ on the sky, arriving at antenna $i$ with delay $\tau_i$ and antenna $j$ with delay $\tau_j$ with respect to the electric field at the origin which we denote as $s(t,{\bf \hat{s}})$. We allow for two different types of reflections: First, we allow reflections within the analogue path of each $i^{th}$ antenna which we denote as $r_i(\tau,{\bf \hat{s}})$. We also allow for single reflections between any $i-j$ antenna pair which we denote as $r_{ij}(\tau',{\bf \hat{s}})$. Our choice of arbitrary $\tau'$, for now, allows for multi-path propogation between antennas, though we expect it to be dominated by the geometrical delay between the antenna pair. The electric field at antenna $i$ is given by
\begin{equation}
s_i(t,{\bf \hat{s}}) = \int d \tau' r_i(\tau',{\bf \hat{s}})s(t+\tau_i-\tau') + \sum_{j \ne i} \int d \tau' s(t + \tau_j - \tau_{ij}) r_{ij}(\tau',{\bf \hat{s}})
\end{equation}  
In an FX correlator, the electric field is sampled, Fourier transformed, and cross multipled between antenna pairs to form visibilities. The Fourier transform step leaves us with 
\begin{equation}
\widetilde{s}_i(f,{\bf \hat{s}}) = \widetilde{s}(f,{\bf \hat{s}}) \left[ \int d \tau' r_i(\tau',{\bf \hat{s}})e^{2 \pi i(\tau_i - \tau')f} + \sum_{j\ne i} \int d \tau' e^{2 \pi i(\tau_j - \tau_{ij})f} r_{ij}(\tau',{\bf \hat{s}}) \right]
\end{equation}
Multiplying and averaging gives us the visibility for the single source we obtain
\begin{align}\label{eq:SingleSource}
v_{ij}'(f,{\bf \hat{s}}) & = \langle \widetilde{s}_i(f,{\bf \widehat{k}}) \widetilde{s}_j(f,{\bf \hat{s}}) \rangle_t \nonumber \\
&=d \Omega  I(f,{\bf \hat{s}}) g_i(f) g_j^*(f) a_i(f,{\bf \hat{s}})a_j^*(f,{\bf \hat{s}}) e^{2 \pi i{\bf u_{ij}} \cdot {\bf \hat{s}}} + d \Omega I(f,{\bf \hat{s}}) \sum_{\ell \ne j} g_i(f)a_i(f,{\bf \hat{s}}) C_{\ell j}^*(f,{\bf \hat{s}}) e^{2 \pi i {\bf u_{\ell i} \cdot {\bf \hat{s}}}} \nonumber \\ 
& + d \Omega I(f,{\bf \hat{s}})\sum_{k \ne i} g_j^*(f) a_j^*(f) C_{ki}(f,{\bf \hat{s}}) e^{2 \pi i {\bf u_{kj}\cdot {\bf \hat{s}}}} + d \Omega I(f,{\bf \hat{s}}) \sum_{k\ne i}\sum_{\ell \ne j} C_{ki}(f,{\bf \hat{s}}) C_{j\ell}^*(f,{\bf \hat{s}}) e^{2 \pi i {\bf u_{k\ell}}\cdot{\bf \hat{s}}},
\end{align}
where $g_i(f) a_i(f,{\bf \hat{s}}) = \int d \tau r_i(\tau,{\bf \hat{s}}) e^{2 \pi i f \tau}$ is the effective direction dependent gain of the system which can be factored into a directoin dependent and direction independent function where $g_i(f)$ is the gain of the analogue signal chain after the radiation has been absorbed by the feed and $a_i(f,{\bf \hat{s}})$ describes the chromatic electric field response of the antenna. The first term in equation~\ref{eq:SingleSource} is an effective visibility with self-reflections. The two cross terms and the last term involve the mixing of visibilities complementary too the $ij$ baseline and have the potential to introduce significant chromatic features since they potentially insert visibilities on much longer basline lengths. Assuming propogation along a single path directly between the antennas, we may write $C_{ik}(f,{\bf \hat{s}})$ as 
\begin{equation}
C_{ki}(f,{\bf \hat{s}}) = a_i(f,{\bf \hat{s}}_{ki}) \frac{1}{r_{ik}}\left[\frac{d \sigma_k}{d \Omega}(f,{\bf \hat{s}},{\bf \hat{s}}_{ik})\right]^{1/2} e^{2 \pi i \tau_{ik} f}
\end{equation}
Where $r_{ik}$ is the distance between antennas $i$ and $k$, $d \sigma_k (f,{\bf \hat{s}},{\bf \hat{s}}_{ij})/ d \Omega $ is the cross-section of the antenna to scatter radiation from the ${\bf \hat{s}}$ direction to the ${\bf \hat{s}}_{ik}$ direction where ${\bf \hat{s}}_{ik}$ is the unit vector in the direction between antenna $i$ and antenna $k$. Integrating over the primary beam, we obtain a full expression on the effect of the foregrounds. 
\begin{align}
V_{ij}' = \int d \Omega v_{ij}'(f,{\bf \hat{s}}) &= g_i(f) g_j(f)^* \int d \Omega A_{ij}(f,{\bf \hat{s}}) I(f,{\bf \hat{s}})e^{2 \pi i f {\bf u_{ij}} \cdot {\bf \hat{s}}} + g_i(f) \sum_{\ell \ne j} \int d \Omega  A_{i\ell j}(f,{\bf \hat{s}}) I(f,{\bf \hat{s}})e^{2 \pi i {\bf u_{\ell i} } \cdot {\bf \hat{s}} } \nonumber \\ 
& + g_j^*(f) \sum_{k \ne i}  \int d \Omega A^*_{jki}(f,{\bf \hat{s}}) I({\bf \hat{s}},f) e^{2 \pi i {\bf u}_{kj} \cdot {\bf \hat{s}} } + \sum_{k \ne i} \sum_{\ell \ne j}\int d \Omega A_{k i \ell j}(f,{\bf \hat{s}}) I(f,{\bf \hat{s}})e^{2 \pi i {\bf u}_{k\ell} \cdot {\bf \hat{s}}}
\end{align}
which is essentially an ad-mixture of many baselines with different effective primary beams. We denote the effective primary beam of the $i,j$ antenna pair as $A_{ij}(f,{\bf \hat{s}})=a_i(f,{\bf \hat{s}})a_j^*(f,{\bf \hat{s}})$, the effective beam from a single reflection correlated with a direct measurement as $A_{i \ell j} = a_i(f,{\bf \hat{s}}) C^*_{\ell j}(f,{\bf \hat{s}})$ and the correlation between entirely reflected terms as experiencing an effective primary beam of $A_{k i \ell j}=C_{ki}(f,{\bf \hat{s}}) C^*_{j \ell}(f,{\bf \hat{s}})$. In this paper, we focus on the reflection terms within a single antenna element. Hence, we ignore all but the first term for now. Any reflection terms occurring downstream of the conversion by the feed from electromagnetic radiation to voltage are lumped into $g_i(f)$ and reflections occurring within the antenna element enter into $a_i(f,{\bf \hat{s}})$. While reflections within the analogue system are a potential source of contamination, HERA's post-feed analogue signal path is designed to keep all reflections under 35\,m, within the wedge.

The focus of this paper and its companions is the reflection properties of the primary antenna element, so we will focus the rest of our discussion here on $a_i(f,{\bf \hat{s}})$. The primary elements of our dish include a feed and backplane suspended over a fourteen meter dish. We assume a set of discrete reflections within the dish, which without loss of generality are assumed to have frequency independent reflection coefficients\footnote{If we allow each coefficient to be frequency dependent, we can expand each frequency dependent term in a Fourier series of frequency independent terms} hence 
\begin{equation}
a_i(f,{\bf \hat{s}}) = \sum_{n} r_n({\bf \hat{s}}) e^{2 \pi i  \tau_n f}
\end{equation}
Assuming all antennas are identical, we have
\begin{equation}
A_{ij}(f,{\bf \hat{s}}) = \sum_m \sum_n r_n({\bf \hat{s}}) r_m^*({\bf \hat{s}}) e^{2 \pi i(\tau_n-\tau_m)f} = \sum_\alpha A_\alpha({\bf \hat{s}}) e^{2 \pi i \tau_\alpha}
\end{equation}
where we have re-indexed $m$ and $n$ under a single greek index $\alpha$ in the second equality. The effect of internal reflections on a visibility is hence
\begin{equation}
V_{ij}'= \sum_\alpha \int d \Omega A_\alpha({\bf \hat{s}}) e^{2 \pi i \tau_\alpha f} e^{2 \pi i {\bf b}_{ij} \cdot {\bf \hat{s}}f/c} I(f,{\bf \hat{s}})
\end{equation}
Taking the delay transform, one obtains
\begin{equation}
\widetilde{V}_{ij}'(\tau) = \sum_\alpha \widetilde{V}_{ij}^\alpha (\tau - \tau_\alpha)
\end{equation}
where 
\begin{equation}
\widetilde{V}^\alpha_{ij}(\tau) = \int d \tau e^{-2 \pi i f \tau} \int d \Omega A_\alpha({\bf \hat{s}}) e^{2 \pi i {\bf b}_{ij} \cdot {\bf \hat{s}} f/c} I(f,{\bf \hat{s}})
\end{equation}
is the usual delay transform of \citet{Parsons:2012}. $\widetilde{V}_{ij}^alpha(\tau)$.
\end{document}
